---
title: "Linear Models"
author: "Emily Schemanske & Nitanshi Mahajan"
date: "4/15/2023"
output: html_document
---

# Hugging Face BERTweet sentiment model

```{r}
# read huggingface data
data_hf = read.csv('/Users/landise/Downloads/lm_huggingface.csv')

# remove date column
data_hf <- subset(data_hf, select = c(adidas_PercentChangeMA, djia_PercentChangeMA, final_score_MA, pn_score_MA))
data_hf <- na.omit(data_hf)

head(data_hf)
```
```{r}
# scale pn_score_MA so all values are between 0 and 1
data_hf$pn_score_MA <- (data_hf$pn_score_MA - min(data_hf$pn_score_MA)) / (max(data_hf$pn_score_MA) - min(data_hf$pn_score_MA))
head(data_hf, 1000)
```

```{r}
# univariate model for reference
base_mod_hf <- lm(adidas_PercentChangeMA ~ djia_PercentChangeMA, data=data_hf)
summary(base_mod_hf)
```

```{r}
# check model with all predictors
hf_full <- lm(adidas_PercentChangeMA ~ ., data=data_hf)
summary(hf_full)
```

```{r}
# delete insignificant predictor
hf_mod <- lm(adidas_PercentChangeMA ~ djia_PercentChangeMA + pn_score_MA, data=data_hf)
summary(hf_mod)
```


# NLTK VADER sentiment model

```{r}
# load nltk scores
d = read.csv('/Users/landise/Downloads/lm.csv')

# remove date column
data <- subset(d, select = c(adidas_PercentChangeMA, djia_PercentChangeMA, nltk_positive_MA, nltk_negative_MA, nltk_neutral_MA, nltk_compound_MA))
head(data)
```

```{r}
# same base model as before
base_mod <- lm(adidas_PercentChangeMA ~ djia_PercentChangeMA, data=data)
summary(base_mod)
```

```{r}
# check full model
nltk_full = lm(adidas_PercentChangeMA ~ ., data=data)
summary(nltk_full)
```

```{r}
# delete most insignificant pred
nltk_less1 = lm(adidas_PercentChangeMA ~ djia_PercentChangeMA + nltk_negative_MA + nltk_neutral_MA + nltk_compound_MA, data=data)
summary(nltk_less1)
```

All predictors are significant.

# Diagnostics

## Plots

```{r}
plot(nltk_less1)
```


```{r}
# lag plot to check for autocorrelation
n <- length(residuals(nltk_full))
plot(tail(residuals(nltk_full), n-1) ~ head(residuals(nltk_full), n-1), xlab=expression(hat(epsilon)[i]), ylab=expression(hat(epsilon)[i+1]))
abline(h=0, v=0, col=grey(0.75))
```
There is strong evidence of autocorrelation - not uncommon in time series.

## Transform the predictors

```{r}
library(car)
durbinWatsonTest(nltk_less1)
```

p-value indiates strong evidence of autocorrelation.

```{r}
# another package for DW test
library(orcutt)
dwtest(nltk_less1)
```

Source: https://rpubs.com/apricitea/handling-autocorrelation

### NLTK final model transformation

```{r}
# get best RHO for transformation
mod_transformed <- cochrane.orcutt(nltk_less1, convergence=5, max.iter=1000)
mod_transformed
```


```{r}
# transform predictors and target
rho <- mod_transformed$rho
y.trans <- data$adidas_PercentChangeMA[-1]-data$adidas_PercentChangeMA[-309]*rho
x.trans <- (data$djia_PercentChangeMA)[-1]-(data$djia_PercentChangeMA)[-309]*rho
x1.trans <- (data$nltk_negative_MA)[-1]-(data$nltk_negative_MA)[-309]*rho
x2.trans <- (data$nltk_neutral_MA)[-1]-(data$nltk_neutral_MA)[-309]*rho
x3.trans <- (data$nltk_compound_MA)[-1]-(data$nltk_compound_MA)[-309]*rho

# model with transformed preds
mod_transformed <- lm(y.trans ~ x.trans + x1.trans + x2.trans + x3.trans)
summary(mod_transformed)
```

```{r}
dwtest(mod_transformed)
```

p-value near 1 indicates we have fixed the issue of autocorrelation

```{r}
n <- length(residuals(mod_transformed))
plot(tail(residuals(mod_transformed), n-1) ~ head(residuals(mod_transformed), n-1), xlab=expression(hat(epsilon)[i]), ylab=expression(hat(epsilon)[i+1]))
abline(h=0, v=0, col=grey(0.75))
```
Plot shows no more evidence of autocorrelation.

## Select final vars

```{r}
summary(mod_transformed)
```

```{r}
# y.trans <- data$adidas_PercentChangeMA[-1]-data$adidas_PercentChangeMA[-309]*rho
# x.trans <- (data$djia_PercentChangeMA)[-1]-(data$djia_PercentChangeMA)[-309]*rho
# x1.trans <- (data$nltk_negative_MA)[-1]-(data$nltk_negative_MA)[-309]*rho
# x2.trans <- (data$nltk_neutral_MA)[-1]-(data$nltk_neutral_MA)[-309]*rho
# x3.trans <- (data$nltk_compound_MA)[-1]-(data$nltk_compound_MA)[-309]*rho

mod_transformed <- lm(y.trans ~ x.trans + x2.trans + x3.trans)
summary(mod_transformed)
```

```{r}
mod_transformed <- lm(y.trans ~ x.trans + x3.trans)
summary(mod_transformed)
```

## Try full model transformation

```{r}
full_transformed <- cochrane.orcutt(nltk_full, convergence=5, max.iter=1000)
full_transformed
```

```{r}
rho <- full_transformed$rho
y.trans <- data$adidas_PercentChangeMA[-1]-data$adidas_PercentChangeMA[-309]*rho
x.trans <- (data$djia_PercentChangeMA)[-1]-(data$djia_PercentChangeMA)[-309]*rho
x1.trans <- (data$nltk_negative_MA)[-1]-(data$nltk_negative_MA)[-309]*rho
x2.trans <- (data$nltk_neutral_MA)[-1]-(data$nltk_neutral_MA)[-309]*rho
x3.trans <- (data$nltk_compound_MA)[-1]-(data$nltk_compound_MA)[-309]*rho
x4.trans <- (data$nltk_positive_MA)[-1]-(data$nltk_positive_MA)[-309]*rho

full_transformed <- lm(y.trans ~ x.trans + x1.trans + x2.trans + x3.trans + x4.trans)
summary(full_transformed)
```

```{r}
full_transformed <- lm(y.trans ~ x.trans + x1.trans + x2.trans + x4.trans)
summary(full_transformed)
```

```{r}
full_transformed <- lm(y.trans ~ x.trans + x1.trans + x4.trans)
summary(full_transformed)
```

```{r}
full_transformed <- lm(y.trans ~ x.trans + x1.trans)
summary(full_transformed)
```

```{r}
plot(full_transformed)
```

```{r}
n <- length(residuals(full_transformed))
plot(tail(residuals(full_transformed), n-1) ~ head(residuals(full_transformed), n-1), xlab=expression(hat(epsilon)[i]), ylab=expression(hat(epsilon)[i+1]))
abline(h=0, v=0, col=grey(0.75))
```

```{r}
base_transformed <- lm(y.trans ~ x.trans)
summary(base_transformed)
```

Both models show simliar results.
